\documentclass[../main.tex]{subfiles}
\begin{document}
\section{Introduction}
\label{sec:introduction}
%---background---
%---macro_problem---
Empirical results in economics are sometimes far more fragile than their standard errors suggest. While the credibility revolution has made remarkable strides in ensuring that estimates recover causal relationships through credible instruments and carefully constructed control groups \parencite{angrist2010credibility, athey2017state}, a persistent mechanical vulnerability remains. Specifically, even a rigorously identified study can collapse upon the removal of a fraction of one percent of its observations \parencite{broderick2020automatic}. This severe sensitivity frequently goes unnoticed because aggregate statistical significance depends on small, bounded subsets of the data whose influence classical diagnostic tools are structurally unable to detect.

%---conceptual_gap---
Classical single-observation diagnostics, such as Cook's distance, DFBETAS and leverage, are designed strictly to flag individual outliers in isolation. This design assumption breaks down when observations exert influence jointly, as masking effects and higher-order interactions allow coordinated subsets to evade detection entirely \parencite{chatterjee1986influential, rousseeuw1990unmasking}. Conversely, while robust estimation techniques prioritize aggregate model stabilization, they inherently obscure rather than locate the exact bounded subsets driving the empirical conclusion.

%---methodological_intervention---
Historically, isolating sets of jointly influential observations remained mathematically intractable due to the combinatorial explosion inherent in evaluating all possible sample subsets. The Most Influential Sets (MIS) framework explicitly resolves this computational barrier \parencite{kuschnig2021hidden}. By reformulating worst-case sensitivity as a linear fractional program, the algorithm guarantees the exact identification of bounded subsets that maximally destabilize aggregate parameters. Yet, exact algorithmic identification alone cannot distinguish genuine structural fragility from expected sampling variation. To establish valid inference, the framework evaluates the statistical significance of these sets through the lens of Extreme Value Theory \parencite{konrad2025testing}. By shifting the analytical focus toward exact set-level influence, this methodology provides a disciplined statistical environment to audit the robustness of econometric claims.

%---scope_and_objectives---
The primary objective of this thesis is to systematically evaluate the empirical utility of the MIS framework in resolving structural vulnerabilities within applied econometrics. The research design proceeds across three analytical domains. First, the study examines the mathematical conditions under which exact set-level identification exposes failures that high-breakdown robust estimators cannot detect. This theoretical baseline is subsequently subjected to extensive Monte Carlo simulations; by incorporating heteroskedasticity, masking, and endogeneity, these experiments validate both the algorithmic recovery of influential subsets and the inferential precision of EVT thresholds. Finally, the framework is deployed to audit contested empirical findings within foundational econometric literature and algorithmic fairness benchmarks. Through this tripartite evaluation, the thesis translates set-level influence from a theoretical abstraction into an operational diagnostic standard.

%---structure---
The remainder of this thesis proceeds as follows. Chapter 2 evaluates the historical development of classical diagnostic tools and robust estimators, establishing the theoretical necessity for set-level inference. Chapter 3 formalizes the proposed methodology, detailing the mathematical mechanics of the MIS algorithm and its associated hypothesis testing framework under Extreme Value Theory. The empirical utility of this framework is subsequently tested in Chapter 4 through a Monte Carlo simulation study that examines algorithmic scalability and accuracy under severe masking and swamping conditions. Chapter 5 applies the methodology to real-world data, auditing foundational econometric datasets and algorithmic fairness benchmarks to detect structural vulnerabilities obscured by aggregate metrics. Finally, Chapter 6 synthesizes the core contributions and outlines trajectories for future research.

\end{document}